---
title: "Exploring Bulgarian Districts and Whether They Exhibit Spatial Spillovers"
author: "Milena Atanasova"
output:
  html_document:
    df_print: paged
---

### Read and explore the data

```{r}
library(spdep); library(sf); library(tmap); library(leaflet)
```

```{r}
bulgaria <- st_read("C:\\Users\\User\\Documents\\UNITN\\Geospatial\\For_GITHUB_project\\Plovdiv_Infographics\\Data\\Climate\\gdf_bg_climateinfo.geojson")
plot(st_geometry(bulgaria))

dim(bulgaria)
names(bulgaria)
str(bulgaria)
head(bulgaria)
```

# Creating spatial weights matrices 

Define the neighbourhood relationships among the spatial units. There are differnet methods 
in which a weight matrix can be defined. Below are shown some examples.

```{r}
# Spatial neighbors
coords <- st_centroid(st_geometry(bulgaria)) #unique reference spatial coordinate
plot(st_geometry(bulgaria), border="blue") 
plot(coords, add = TRUE)

# For Sofia, as its shape is particular, the centroid is very close to the border with the region of Sofia capital.
```

## Defining neighbors

#### 1.KNN
```{r}
knn1BG <- knn2nb(knearneigh(coords,k=1))
plot(st_geometry(bulgaria), border="grey") 
plot(knn1BG, coords, add=TRUE)
```

#### 2.Critical cut-off neighborhood

```{r}
#minimum threshold distance
knn1IT <- knn2nb(knearneigh(coords,k=1))
all.linkedT <- max(unlist(nbdists(knn1IT, coords))) 
all.linkedT

dnb75 <- dnearneigh(coords, 0, 75); dnb75
dnb100 <- dnearneigh(coords, 0, 100); dnb100

plot(st_geometry(bulgaria), border="grey")
plot(dnb75, coords, add=TRUE, col="blue")
plot(dnb100, coords, add=TRUE, col="red")

```

#### 3.Contiguity-based neighbourhood

```{r}
# if share a common boundary

contnb_q <- poly2nb(bulgaria, queen=T)
contnb_q
plot(st_geometry(bulgaria), border="grey") 
plot(contnb_q, coords, add=TRUE)

```


### Defining spatial weights

```{r}
dnb75.listw <- nb2listw(dnb75,style="W") #form of a list
dnb100.listw <- nb2listw(dnb100,style="W")
listw2mat(dnb75.listw)
```

### Building free-form spatial weight matrices 

```{r}
#weights as inverse functions of the distance among centroids
distM <- st_distance(coords)/1000
class(distM) <- "matrix" #distance matrix
distM
# Three possible weight matrices
#using inverse functions of the matrices
W1 <- 1/(1+distM); diag(W1) <- 0
W2 <- 1/(1+distM^2); diag(W2) <- 0
W3 <- exp(0.1*-distM);diag(W3) <- 0

#Row-standardize them 
W1s <- W1/rowSums(W1) 
W2s <- W2/rowSums(W2) 
W3s <- W3/rowSums(W3) 
#convert
listW1s <- mat2listw(W1s)
listW2s <- mat2listw(W2s)
listW3s <- mat2listw(W3s)
```


# The Moran's I test of spatial autocorrelation 
(a global measure)

First, let's merge data from the Bulgarian Statistical Institute as well.
```{r}
library(readxl)
bgstats <- read_excel("C:\\Users\\User\\Documents\\UNITN\\Geospatial\\For_GITHUB_project\\R_SpatialRegression\\BG_statistics.xlsx")
colnames(bgstats)[1] = "NAME_LATN"
prov <- merge(bulgaria, bgstats, by="NAME_LATN")
head(prov)
```

Below are some plots of several variables.

```{r}

plot(prov[c("HealthEstablishments")], main="Health Establishments") 

tm_shape(prov) + 
  tm_polygons("AccommodationEstablishments", title = "Accommodation Establishments") +
  tm_layout(legend.outside = TRUE) 

tm_shape(prov) + 
  tm_polygons("temperature_average", title = "Temperature") +
  tm_layout(legend.outside = TRUE) 

require(RColorBrewer)
qpal <- colorQuantile("OrRd", prov$GDPmlnbg, n=9) 
leaflet(prov) %>%
  addPolygons(stroke = FALSE, fillOpacity = .8, smoothFactor = 0.2, color = ~qpal(GDPmlnbg)) %>%
  addTiles()

```


## Moran's test

Trying different variables, but it appears that p-values are quite high, so it seems that Spatial autocorrelation of the variables is not statistically significant.

```{r}
# GDP
moran.mc(prov$GDPmlnbg, dnb75.listw, nsim = 999) 

#Temperature, try a different matrix
moran.test(prov$temperature_average, dnb100.listw, randomisation = TRUE)
moran.mc(prov$temperature_average, listW1s, nsim=999)


moran.test(prov$precipitation_average, dnb100.listw, randomisation = TRUE)
moran.mc(prov$precipitation_average, listW1s, nsim=999)

moran.test(prov$GeneratedMunicipalWaste, dnb100.listw, randomisation = TRUE)
moran.mc(prov$GeneratedMunicipalWaste, listW1s, nsim=999)

moran.mc(prov$wind_speed_average, listW1s, nsim=999)

```



### The Moran's I test of spatial autocorrelation in OLS residuals of a linear regression model

```{r}
Lm1 <- lm(temperature_average ~ precipitation_average + wind_speed_average, prov)
summary(Lm1) 

#plot of the studentized residuals
prov$studres <- rstudent(Lm1)
qpal <- colorQuantile("OrRd", prov$studres, n=4) 
leaflet(prov) %>%
  addPolygons(stroke = FALSE, fillOpacity = .8, smoothFactor = 0.2, color = ~qpal(studres)) %>%
  addTiles() 

#There can be seen some similarities in the residuals in some regions which are close next to each other.

```

```{r}
#Applying the Moran's I test in the OLS residuals 
lm.morantest(Lm1,dnb100.listw,resfun=rstudent)
lm.morantest(Lm1,listW1s,resfun=rstudent)
```

Again not really significant even though a bit lower p-values.


```{r}
library(boot)
Lm1.lmx <- lm(temperature_average ~ precipitation_average + wind_speed_average, data = prov, x = TRUE)
MoraneI.boot <- function(var, i, ...) {
  var <- var[i]
  lmres <- lm(var ~ Lm1.lmx$x - 1)
  return(moran(x=residuals(lmres), ...)$I)	  
}
boot1 <- boot(residuals(Lm1.lmx),statistic=MoraneI.boot,  
                R=999, sim="permutation", listw=listW1s,  
                n=length(listW1s$neighbours), S0=Szero(listW1s))
ti <- (boot1$t0 - mean(boot1$t))/sqrt(var(boot1$t))  
boot1
plot(boot1)
# We can see there is some pattern in the residuals, they don't follow the quantiles excatly
#But still not too far off
```


# Local Moran I and test
(local spatial clusters)

## 1. Moran scatterplot 

```{r}
#GDP
mplot <- moran.plot(prov$GDPmlnbg, listw=dnb75.listw, main="Moran scatterplot GDP")
# Since the pattern of Sofia is particular it is way far off, it is difficult to make inference

mplot <- moran.plot(prov$temperature_average, listw=listW1s, main="Moran scatterplot Temperature")
grid()
```

From the plot some things could be pointed out:
There seem to be some division in the four quadrants;
Higher temperature seems to be associated with lower local average.
Sofia and Blagoevgrad exert the most influence on the regression line.

```{r}
prov$hat_value <- mplot$hat
tm_shape(prov) + tm_polygons("hat_value") 
```

Below, the most influencial units.

```{r}

mplot <- moran.plot(prov$temperature_average, listw=listW1s, main="Moran scatterplot", 
         return_df=F)
hotspot <- as.numeric(row.names(as.data.frame(summary(mplot))))

prov$wx <- lag.listw(listW1s, prov$temperature_average)

prov$quadrant <- rep("None", length(prov$temperature_average))
for(i in 1:length(hotspot))  {
  if (prov$temperature_average[hotspot[i]]>mean(prov$temperature_average) & prov$wx[hotspot[i]]> mean(prov$wx)) 
        prov$quadrant[hotspot[i]] <- "HH" 
  if (prov$temperature_average[hotspot[i]]>mean(prov$temperature_average) & prov$wx[hotspot[i]]< mean(prov$wx)) 
        prov$quadrant[hotspot[i]] <- "HL" 
  if (prov$temperature_average[hotspot[i]]<mean(prov$temperature_average) & prov$wx[hotspot[i]]<mean(prov$wx)) 
        prov$quadrant[hotspot[i]] <- "LL" 
  if (prov$temperature_average[hotspot[i]]<mean(prov$temperature_average) & prov$wx[hotspot[i]]>mean(prov$wx)) 
        prov$quadrant[hotspot[i]] <- "LH" 
  }
table(prov$quadrant)

#which allows us to plot the map of the regions with influence by typing
tm_shape(prov) + tm_polygons("quadrant")
```


# 2. The Local Moran's I values

```{r}
lmI <- localmoran(prov$temperature_average, listW1s)
head(lmI)
prov$lmI <- lmI[,1]
tm_shape(prov) + 
    tm_polygons("lmI", title = "Local Moran's I values") 

# We can see some of the local regions that are being identified
```

```{r}
lmIp1 <- localmoran_perm(prov$temperature_average, listW1s, nsim = 9999, iseed = 1) 
lmIp2 <- localmoran_perm(prov$GDPmlnbg, listW1s, nsim = 9999, iseed = 1)

#Things do not seem significant.
```



# Spatial regression

Let's see if there exists the need to include spatial characteristics when
computing models.

```{r}
library(spatialreg)
```

```{r}
#The Lagrange multiplier (LM) test of spatial dependence on OLS residuals

Lm1 <- lm(temperature_average ~ precipitation_average + wind_speed_average, prov)
natOLSlmTests <- lm.LMtests(Lm1, dnb100.listw, 
                    test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
summary(natOLSlmTests)

```

After having tried with different weight matrices, it seems there is no presence of
importance of the spatial distribution for the model about average temperature.


```{r}
#Also for the GDP not preferred any other models which consider the spatial autocorrelation

Lm2 <- lm(GDPmlnbg ~ Area + PublicWaterSupply + GeneratedMunicipalWaste + AccommodationEstablishments, prov)
natOLSlmTests <- lm.LMtests(Lm2, dnb100.listw, 
                    test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
summary(natOLSlmTests)

```

From the analysis conducted, it appears that overall in Bulgarian districts, 
the geographical location does not play a big role for some statistical indicators.
The local/global effects cannot be seen, and in fact since the main and most developed
cities of the country are more or less scattered around the country, it makes sense.

Therefore, if regression is to be done on some factor concerning the variables at hand,
it can be relied on linear regression models with OLS estimator.

There could be some other indicator variables for the country, which could possibly 
possess the need to be analysed with the use of a spatial regression model.

In fact, maybe if analysis is done based on municipalities, some more useful and informative
results could be obtained. However, not much data of this kind is available.


